ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel

### 2025-08-14 paper/EPB-Final-Submission.tex
Final version for EPB publication

\runninghead{}

\author{Anon1\affilnum{}, Anon2\affilnum{}, Anon3\affilnum{}}

\affiliation{\affilnum{}{}}

\begin{abstract}
This paper describes \{ActiveCA\}, an open data product with Canadian
time use data. \{ActiveCA\} is an R data package that contains
analysis-ready data related to active travel spanning almost 40 years,
extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
(2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) cicles
from the General Social Survey (GSS). Active travel is characterized by
mode, with walking being part of every cycle and bicycling starting in
1992. The attributes of active trips are the types of locations of
origins and destinations, the duration of trips, and episode weights for
expanding the trips to population-wide estimates. Based on the year of
the survey, a variety of locations are coded. In earlier cycles, these
include home, work or school, and other's home, whereas in later cycles
these are augmented with locations such as grocery stores, restaurants,
outdoor destinations, and others. The geographical resolution includes
the province and whether the episode was in an urban or rural setting.
\end{abstract}

\keywords{Active; mobility; walking; cycling; travel time; time-use;
General Social Survey}

@@

\{ActiveCA\} provides analysis-ready data concerning active travel in
Canada spanning a period of almost 40 years, obtained from the TUS
cycles of the Canadian GSS. The GSS program is designed to provide
cross-sectional data on topics of interest to improve the well-being of
Canadians. As part of this program, every five to seven years the survey
is done on the topic of time use (TUS). Concretely, \{ActiveCA\} covers
Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015),
and 34 (2022) of the TUS. Time use data in these surveys is coded using
a very fine grain, from time spent in chores, leisure, and sleeping, to
time spent working or at school. These surveys have proved valuable in
investigations of mobility and quality of life
\citep{spinneyTransport2009}, the relationship between active travel and
transit use \citep{lachapelleLonger2016}, and travel behavior and time
poverty \citep{kimFacing2024}, to name but a few examples.

Using Public Use Microdata Files (PUMFs) from the Time Use Survey (TUS),
we extracted all data necessary to characterize active travel in Canada
- specifically, episodes in which the activity involved moving between
an origin and a destination by walking or cycling. Although Statistics
Canada provides PUMFs and accompanying documentation for the GSS program
\citep[see][]{statisticscanada2024}, accessing and preparing these files
for analysis is not a straightforward task due to their size and
complexity. The process of extracting information of interest from the
source files is time-consuming, tedious, and challenging and/or prone to
error due to the expertise required to work with these files. To create
\{ActiveCA\} we selected, labelled, and processed the TUS cycles to make
them ready for analysis.

\{ActiveCA\} is distributed as an \texttt{R} package with a number of
data objects and their documentation. \texttt{R} packages contain code,
data, and documentation in a standardized format that can be installed
by R users via a software repository, such as CRAN (Comprehensive R
Archive Network) or GitHub, which makes them an adroit medium to
distribute analysis-ready data.

@@

The rest of this paper discusses the sources of data, and the process
implemented to retrieve and package them. Then, we explain how to use
the package and show some selected examples of analysis to whet the
imagination of potential users. This ODP provides not only data that are
easy to use, but also all the code and documentation that make this a
reproducible research project. In summary, \{ActiveCA\} aims to
implement and inspire the best principles of open spatial sciences
\citep{paez_open_2021, brunsdon_opening_2021}.

@@

Most respondents to the 2022 TUS completed it online, reflecting
Statistics Canada's effort to adapt to technological changes and growing
time demands by offering greater flexibility and convenience
\citep{statisticscanada2022}. While such methodological changes may
affect data comparability over time, it is not possible to determine
whether observed differences result from actual population changes or
shifts in data collection methods. Despite rigorous efforts to ensure
data quality, the use of electronic questionnaires may have influenced
estimates. Statistics Canada assessed the impact of collection mode on a
limited set of key questions, constrained by sample size. Importantly,
none of the variables used in this research are \texttt{R} package in
the 2022 PUMF User Guide as unsuitable for trend analysis.

@@

The survey encompasses both rural and urban areas, including
metropolitan and non-metropolitan regions, to ensure a diverse and
representative sample. For sampling, the ten provinces were divided into
geographic strata. Several Census Metropolitan Areas (CMAs) - such as
St.~John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa,
Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and Vancouver
- were treated as separate strata. Additional strata grouped other CMAs
within Quebec, Ontario, and British Columbia, as well as non-CMA areas
within each province.

The Public Use Microdata are released by Statistics Canada in two files:
a \emph{Main file} and an \emph{Episode file}. The files are linked by
keys that identify households, individuals, and episodes (i.e.,
activities) conducted by individuals. We discuss these files in more
detail in the following section.

@@

The \emph{Main file} of the time use survey compiles a large array of
aggregated data, summarizing the answers to the questionnaire that
describe households and individuals, as well as derived variables that
summarize the respondents' use of time use across different activities,
locations, and social interactions. This file documents the time and
duration that respondents allocate to each activity and location. The
\emph{Main file} provides a overview of daily routines and social
dynamics, not focusing on individual activity episodes. Additionally,
this file categorizes activities into bigger groups and subcategories,
facilitating the data's analytical utility with additional metrics such
as total transit time, time spent with household members, and counts of
activities and episodes.

Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first
six variables of the TUS PUMF 2015 \emph{Main file} (Cycle 29). Each row
in the table correspond to a survey respondent, while the columns refer
the following information: record identification (\texttt{PUMFID}), the
person's weight (\texttt{WGHT\_PER}), the month the survey data was
collected (\texttt{SURVMNTH}), the respondent's age group
(\texttt{AGEGR10}), the respondent's sex (\texttt{SEX}), and the
respondent's marital status (\texttt{MARSTAT}).

The \emph{Main file} of the 2015 GSS surveys includes a total of 17,390
respondents, representing 29,766,399 individuals and 848 variables. For
discrete variables, Statistics Canada has assigned specific codes to the
possible values, with each code accompanied by a label. For instance, in
the case of the variable \texttt{SURVMNTH}, a value of \texttt{1} means
\texttt{January\ 2016}, \texttt{2} means \texttt{February\ 2016},
\texttt{3} corresponds to \texttt{March\ 2016}, and so on.

@@

\begingroup\fontsize{8}{10}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item Legend: PUMFID: record identification. WGHT\_PER:  person weight. SURVMNTH: survey month of data collection. AGEGR10: age group of the respondent. SEX: sex of the respondent. MARSTAT: marital status of the respondent.
\end{TableNotes}
\begin{longtable}[t]{cccccc}
\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main file of the 2015 GSS.}\\
\toprule
PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
\midrule
10000 & 616.6740 & 7 & 5 & 1 & 5\\
10001 & 8516.6140 & 7 & 5 & 1 & 1\\
10002 & 371.7520 & 1 & 4 & 2 & 1\\
10003 & 1019.3135 & 3 & 6 & 2 & 5\\
10004 & 1916.0708 & 9 & 2 & 1 & 6\\
\addlinespace
10005 & 1952.2015 & 4 & 1 & 1 & 6\\
10006 & 5761.5528 & 8 & 1 & 1 & 6\\
10007 & 466.0426 & 6 & 5 & 2 & 3\\
10008 & 2479.2991 & 2 & 2 & 2 & 1\\
10009 & 1436.1641 & 8 & 6 & 1 & 3\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

@@

The episode is a much bigger file that records detailed data for each
activity episode reported by respondents. Each episode represents a
single activity and its duration, and the sum of all episodes throughout
the day adds up to 24 hours. Each entry in this file includes the start
and end times of the activity, the duration, location, and accompanying
social context, informing when and where activities occurred and with
whom. The focus of the \emph{Episode file} is not on the characteristics
of the respondents, but on the characteristics of the activities, and
the data are structured around the numerous activity instances that
compose a day of the respondent. Although respondent-specific
characteristics are not included within the \emph{Episode file}, it is
possible to link the \emph{Main file} and the \emph{Episode file} by
using a key present in both the \emph{Main} and \emph{Episode files}.

Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an
example of the \emph{Main file} structure, Table
\ref{tab:ep-2015-unprocessed} presents the first seven episodes for the
record identification number 10041 and some variables from the TUS PUMF
2015 \emph{Episode file} (Cycle 29). Each row in the table corresponds
to an episode associated with the specified record identification
(\texttt{PUMFID} = \texttt{10041}), including the episode's weight
(\texttt{WGHT\_EPI}), episode number (\texttt{EPINO}), activity code
(\texttt{TUI\_01}), episode duration (\texttt{DURATION}), and episode
location (\texttt{LOCATION}).

In total, the \emph{Episode file} of the 2015 GSS surveys contains
274,108 records, representing 461,837,622 episodes and 527 variables
detailing the episodes. Similar to the \emph{Main file}, Statistics
Canada has created codes for the discrete variables, with each value
corresponding to a label.

@@

\begingroup\fontsize{8}{10}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item Legend: PUMFID: record identification. EPINO: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
\end{TableNotes}
\begin{longtable}[t]{cccccc}
\caption{\label{tab:gss-epi-file-2015}\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number 10041.}\\
\toprule
PUMFID & EPINO & WGHT\_EPI & TUI\_01 & DURATION & LOCATION\\
\midrule
10041 & 1 & 1353.818 & 1 & 210 & 300\\
10041 & 2 & 1353.818 & 2 & 40 & 300\\
10041 & 3 & 1353.818 & 27 & 15 & 300\\
10041 & 4 & 1353.818 & 7 & 15 & 315\\
10041 & 5 & 1353.818 & 9 & 180 & 300\\
\addlinespace
10041 & 6 & 1353.818 & 6 & 15 & 300\\
10041 & 7 & 1353.818 & 18 & 120 & 300\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

@@

Figure \ref{fig:process-figure} presents a diagram illustrating the
processes applied to the \emph{Main} and \emph{Episode files} to create
the \{ActiveCA\} datasets. For each TUS cycle, we reviewed the
\emph{Episode files} to identify movement episodes involving walking or
cycling. This allowed us to also select the activities immediately
before and after the movement episode, helping to infer the trip's
purpose and determine its origin and destination. Active trips were
identified by their corresponding activity codes, accounting for
variations across survey cycles.

@@

For the \emph{Main files}, we selected socioeconomic variables to help
profile individuals engaged in active travel episodes. These included
key indicators such as age group, sex, marital status, and number of
children, among others. As with the \emph{Episode files}, we labeled and
factored the coded variables. Using the appropriate identifiers, we
joined the Episode and \emph{Main files} to determine the province and
the type of population centre - whether it is a CMA, a Census
Agglomeration (CA), or a non-CMA/CA area.

@@

This section presents some potential applications of the \{ActiveCA\}
\texttt{R} package. In fact, we expect that the application of this
package to extend beyond our pre-imagined range of uses. The
installation instruction and also some examples of application of the
\{ActiveCA\} \texttt{R} package are available in the \texttt{vignettes},
available in the Github repository.

@@

Table \ref{tab:processed-obs} displays the total number of records
processed for Main and \emph{Episode files}. For the \emph{Main files},
a total of 101,667 records were processed, referring to all records from
the Time Use Surveys from 1986 to 2022, that together represents more of
181,526,641 respondents. It also presents the total cases of active
trips episodes identified. In total 23,513 records with register of
active travel activity. Together, these records account for 44,316,110
episodes.

\begingroup\fontsize{8}{10}\selectfont

\begin{longtable}[t]{rr>{}r|rr}
\caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of records processed.}\\
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Main} & \multicolumn{2}{c}{Episode} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Year & Count & Weighted & Count\_ep & Weighted\_ep\\
\midrule
2022 & 12336 & 32136802 & 1765 & 6041032\\
2015 & 17390 & 29766399 & 3496 & 6634387\\
2010 & 15390 & 28075610 & 4615 & 8516753\\
2005 & 19597 & 26095819 & 5866 & 7583838\\
1998 & 10749 & 24260137 & 1789 & 3606987\\
\addlinespace
1992 & 9815 & 21294313 & 1635 & 3691918\\
1986 & 16390 & 19897562 & 4347 & 8241196\\
\bottomrule
\end{longtable}
\endgroup{}

Table \ref{tab:main-2015-processed} shows the first ten rows and first
six variables of the TUS PUMF 2015 Main File (Cycle 29), displayed in
\ref{tab:main-2015-unprocessed} before our processing. Table
\ref{tab:ep-2015-processed} presents the walking episodes for the record
identification number \texttt{10041} from the TUS PUMF 2015
\emph{Episode file} (Cycle 29), previously displayed in Table
\ref{tab:ep-2015-unprocessed}. Only the unique active travel episode
appear in Table \ref{tab:ep-2015-processed} since the records were
filtered to select cases with walking or cycling episodes. For both
cases, Tables \ref{tab:main-2015-processed} and
\ref{tab:ep-2015-processed} contain labeled variables, facilitating the
interpretation of the data.

\begingroup\fontsize{8}{10}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item Legend: PUMFID: record identification. WGHT\_PER:  person weight. SURVMNTH: survey month of data collection. AGEGR10: age group of the respondent. SEX: sex of the respondent. MARSTAT: marital status of the respondent.
\end{TableNotes}
\begin{longtable}[t]{cccccc}
\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the 2015 TUS Main File.}\\
\toprule
PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
\midrule
10000 & 616.6740 & July & 55 to 64 years & Male & Divorced\\
10001 & 8516.6140 & July & 55 to 64 years & Male & Married\\
10002 & 371.7520 & January & 45 to 54 years & Female & Married\\
10003 & 1019.3135 & March & 65 to 74 years & Female & Divorced\\
10004 & 1916.0708 & September & 25 to 34 years & Male & Single, never married\\
\addlinespace
10005 & 1952.2015 & April & 15 to 24 years & Male & Single, never married\\
10006 & 5761.5528 & August & 15 to 24 years & Male & Single, never married\\
10007 & 466.0426 & June & 55 to 64 years & Female & Widowed\\
10008 & 2479.2991 & February & 25 to 34 years & Female & Married\\
10009 & 1436.1641 & August & 65 to 74 years & Male & Widowed\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

\begingroup\fontsize{8}{10}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item Legend: PUMFID: record identification. EPINO: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
\end{TableNotes}
\begin{longtable}[t]{rrlrlll}
\caption{\label{tab:gss-processed-file-2015}\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number 10041 of the 2015 GSS survey.}\\
\toprule
PUMFID & WGHT\_EPI & Activity & Duration & Origin & Destination & Mode\\
\midrule
10041 & 1353.818 & Transport to or from activity & 15 & Home & Home & Walking\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

@@

\begingroup\fontsize{8}{10}\selectfont

>>>>

\runninghead{Dos Santos \emph{et al}.}

\author{Bruno Dias dos Santos*\affilnum{1}, Mahdis
Moghadasi\affilnum{1}, Antonio Páez\affilnum{1}}

\affiliation{\affilnum{1}{McMaster University}}

\corrauth{Bruno Dias dos Santos}

\email{\href{mailto:dossanb@mcmaster.ca}{\nolinkurl{dossanb@mcmaster.ca}}}

\begin{abstract}
This paper describes \{ActiveCA\}, an open data product with Canadian
time use data. \{ActiveCA\} is an R data package that contains
analysis-ready data related to active travel spanning almost 40 years,
extracted from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
(2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) from the
General Social Survey (GSS). Active travel episodes are characterized by
mode, with walking being part of every cycle and bicycling starting in
1992. The attributes of active trips are the types of locations of
origins and destinations, the duration of trips, and episode weights for
expanding the trips to population-wide estimates. Based on the year of
the survey, a variety of locations are coded. In earlier cycles, these
include home, work or school, and other's home, whereas in later cycles
these are augmented with locations such as grocery stores, restaurants,
outdoor destinations, and others. The geographical resolution includes
the province and whether the episode was in an urban or rural setting.
\end{abstract}

\keywords{Active; mobility; walking; cycling; travel time; time-use;}

@@

\{ActiveCA\} provides analysis-ready data concerning active travel in
Canada spanning a period of almost 40 years, obtained from the TUS
cycles of the Canadian GSS. The GSS program is designed to provide
cross-sectional data on topics of interest to improve the well-being of
Canadians. As part of this program, every five to seven years the survey
is done on the topic of time use. Concretely, \{ActiveCA\} covers Cycles
2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34
(2022) of the TUS. Time use data in these surveys is coded using a very
fine grain, from time spent in chores, leisure, and sleeping, to time
spent working or at school. These surveys have proved valuable in
investigations of mobility and quality of life
\citep{spinneyTransport2009}, the relationship between active travel and
transit use \citep{lachapelleLonger2016}, and travel behavior and time
poverty \citep{kimFacing2024}, to name but a few examples.

Using Public Use Microdata Files (PUMFs) from the TUS, we extracted all
data necessary to characterize active travel in Canada - specifically,
episodes in which the activity involved moving between an origin and a
destination by walking or cycling. Although Statistics Canada provides
PUMFs and accompanying documentation for the GSS program
\citep[see][]{statisticscanada2024}, accessing and preparing these files
for analysis is not a straightforward task due to their size and
complexity. The process of extracting information of interest from the
source files is time-consuming, tedious, and challenging and/or prone to
error due to the expertise required to work with these files. To create
\{ActiveCA\} we selected, labelled, and processed the TUS cycles to make
them ready for analysis.

\{ActiveCA\} is distributed as an R package with a number of data
objects and their documentation. R packages contain code, data, and
documentation in a standardized format that can be installed by R users
via a software repository, such as CRAN (Comprehensive R Archive
Network) or GitHub, which makes them an adroit medium to distribute
analysis-ready data.

@@

The rest of this paper discusses the sources of data, and the process
implemented to retrieve and package them. Then, we show some selected
examples of analysis to whet the imagination of potential users. This
ODP provides not only data that are easy to use, but also all the code
and documentation that make this a reproducible research project. In
summary, \{ActiveCA\} aims to implement and inspire the best principles
of open spatial sciences \citep{paez_open_2021, brunsdon_opening_2021}.

@@

Most respondents to the 2022 TUS completed it online, reflecting
Statistics Canada's effort to adapt to technological changes and growing
time demands by offering greater flexibility and convenience
\citep{statisticscanada2022}. While such methodological changes may
affect data comparability over time, it is not possible to determine
whether observed differences result from actual population changes or
shifts in data collection methods. Despite rigorous efforts to ensure
data quality, the use of electronic questionnaires may have influenced
estimates. Statistics Canada assessed the impact of collection mode on a
limited set of key questions, constrained by sample size. Importantly,
none of the variables used in this research are in the 2022 PUMF User
Guide as unsuitable for trend analysis.

@@

The survey encompasses both rural and urban areas, including
metropolitan and non-metropolitan regions, to ensure a diverse and
representative sample. For sampling, the ten provinces were divided into
geographic stratas. Several Census Metropolitan Areas (CMAs) - such as
St.~John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa,
Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and Vancouver
- were treated as separate strata. Additional strata grouped other CMAs
within Quebec, Ontario, and British Columbia, as well as non-CMA areas
within each province.

The Public Use Microdata are released by Statistics Canada in two files:
a main file and an episode file. The files are linked by keys that
identify households, individuals, and episodes (i.e., activities)
conducted by individuals. We discuss these files in more detail in the
following section.

@@

The main file of the time use survey compiles a large array of
aggregated data, summarizing the answers to the questionnaire that
describe households and individuals, as well as derived variables that
summarize the respondents' use of time use across different activities,
locations, and social interactions. This file documents the time and
duration that respondents allocate to each activity and location. The
main file provides a overview of daily routines and social dynamics, not
focusing on individual activity episodes. Additionally, this file
categorizes activities into bigger groups and subcategories,
facilitating the data's analytical utility with additional metrics such
as total transit time, time spent with household members, and counts of
activities and episodes.

Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first
six variables of the TUS PUMF 2015 main file (Cycle 29). Each row in the
table correspond to a survey respondent, while the columns refer the
following information: record identification (\texttt{PUMFID}), the
person's weight (\texttt{WGHT\_PER}), the month the survey data was
collected (\texttt{SURVMNTH}), the respondent's age group
(\texttt{AGEGR10}), the respondent's sex (\texttt{SEX}), and the
respondent's marital status (\texttt{MARSTAT}).

The main file of the 2015 GSS surveys includes a total of 17,390
respondents, representing 29,766,399 individuals and 848 variables. For
discrete variables, Statistics Canada has assigned specific codes to the
possible values, with each code accompanied by a label. For instance, in
the case of the variable \texttt{SURVMNTH}, a value of \texttt{1} means
\texttt{January\ 2016}, \texttt{2} means \texttt{February\ 2016},
\texttt{3} corresponds to \texttt{March\ 2016}, and so on.

@@

\begin{table}
\centering
\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main file of the 2015 GSS.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{8}{10}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{cccccc}
\toprule
PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
\midrule
10000 & 616.6740 & 7 & 5 & 1 & 5\\
10001 & 8516.6140 & 7 & 5 & 1 & 1\\
10002 & 371.7520 & 1 & 4 & 2 & 1\\
10003 & 1019.3135 & 3 & 6 & 2 & 5\\
10004 & 1916.0708 & 9 & 2 & 1 & 6\\
\addlinespace
10005 & 1952.2015 & 4 & 1 & 1 & 6\\
10006 & 5761.5528 & 8 & 1 & 1 & 6\\
10007 & 466.0426 & 6 & 5 & 2 & 3\\
10008 & 2479.2991 & 2 & 2 & 2 & 1\\
10009 & 1436.1641 & 8 & 6 & 1 & 3\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item Legend: PUMFID: record identification. WGHT\_PER: person weight. SURVMNTH: survey month of data collection. AGEGR10: age group of the respondent. SEX: sex of the respondent. MARSTAT: marital status of the respondent.
\end{tablenotes}
\end{threeparttable}}
\end{table}

@@

The episode is a much bigger file that records detailed data for each
activity episode reported by respondents. Each episode represents a
single activity and its duration, and the sum of all episodes throughout
the day adds up to 24 hours. Each entry in this file includes the start
and end times of the activity, the duration, location, and accompanying
social context, informing when and where activities occurred and with
whom. The focus of the episode file is not on the characteristics of the
respondents, but on the characteristics of the activities, and the data
are structured around the numerous activity instances that compose a day
of the respondent. Although respondent-specific characteristics are not
included within the episode file, it is possible to link the main file
and the episode file by using a key present in both the main and
episodes files.

Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an
example of the main file structure, Table \ref{tab:ep-2015-unprocessed}
presents the first seven episodes for the record identification number
10041 and some variables from the TUS PUMF 2015 episode file (Cycle 29).
Each row in the table corresponds to an episode associated with the
specified record identification (\texttt{PUMFID} = \texttt{10041}),
including the episode's weight (\texttt{WGHT\_EPI}), episode number
(\texttt{EPINO}), activity code (\texttt{TUI\_01}), episode duration
(\texttt{DURATION}), and episode location (\texttt{LOCATION}).

In total, the episode file of the 2015 GSS surveys contains 274,108
records, representing 461,837,622 episodes and 527 variables detailing
the episodes. Similar to the main file, Statistics Canada has created
codes for the discrete variables, with each value corresponding to a
label.

@@

\begin{table}
\centering
\caption{\label{tab:gss-epi-file-2015}\label{tab:ep-2015-unprocessed}Visualization of the  first seven episodes of the record number 10041.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{8}{10}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{cccccc}
\toprule
PUMFID & EPINO & WGHT\_EPI & TUI\_01 & DURATION & LOCATION\\
\midrule
10041 & 1 & 1353.818 & 1 & 210 & 300\\
10041 & 2 & 1353.818 & 2 & 40 & 300\\
10041 & 3 & 1353.818 & 27 & 15 & 300\\
10041 & 4 & 1353.818 & 7 & 15 & 315\\
10041 & 5 & 1353.818 & 9 & 180 & 300\\
\addlinespace
10041 & 6 & 1353.818 & 6 & 15 & 300\\
10041 & 7 & 1353.818 & 18 & 120 & 300\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item Legend: PUMFID: record identification. EPINO: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
\end{tablenotes}
\end{threeparttable}}
\end{table}

@@

Figure \ref{fig:process-figure} presents a diagram illustrating the
processes applied to the main and episode files to create the
\{ActiveCA\} datasets. For each TUS cycle, we reviewed the episode files
to identify movement episodes involving walking or cycling. This allowed
us to also select the activities immediately before and after the
movement episode, helping to infer the trip's purpose and determine its
origin and destination. Active trips were identified by their
corresponding activity codes, accounting for variations across survey
cycles.

@@

For the main files, we selected socioeconomic variables to help profile
individuals engaged in active travel episodes. These included key
indicators such as age group, sex, marital status, and number of
children, among others. As with the episode files, we labeled and
factored the coded variables. Using the appropriate identifiers, we
joined the episode and main files to determine the province and the type
of population centre - whether it is a CMA, a Census Agglomeration (CA),
or a non-CMA/CA area.

@@

This section presents some potential applications of the \{ActiveCA\} R
package. In fact, we expect that the application of this package to
extend beyond our pre-imagined range of uses. The installation
instruction and also some examples of application of the \{ActiveCA\} R
package are available in the vignettes, available in the Github
repository.

@@

Table \ref{tab:processed-obs} displays the total number of records
processed for Main and episode files. For the main files, a total of
101,667 records were processed, referring to all records from the Time
Use Surveys from 1986 to 2022, that together represents more of
181,526,641 respondents. It also presents the total cases of active
trips episodes identified. In total 23,513 records with register of
active travel activity. Together, these records account for 44,316,110
episodes.

\begin{table}
\centering
\caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of records processed.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{rr>{}r|rr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Main} & \multicolumn{2}{c}{Episode} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Year & Count & Weighted & Count\_ep & Weighted\_ep\\
\midrule
2022 & 12336 & 32136802 & 1765 & 6041032\\
2015 & 17390 & 29766399 & 3496 & 6634387\\
2010 & 15390 & 28075610 & 4615 & 8516753\\
2005 & 19597 & 26095819 & 5866 & 7583838\\
1998 & 10749 & 24260137 & 1789 & 3606987\\
\addlinespace
1992 & 9815 & 21294313 & 1635 & 3691918\\
1986 & 16390 & 19897562 & 4347 & 8241196\\
\bottomrule
\end{tabular}
\end{table}

Table \ref{tab:main-2015-processed} shows the first ten rows and first
six variables of the TUS PUMF 2015 Main File (Cycle 29), displayed in
\ref{tab:main-2015-unprocessed} before our processing. Table
\ref{tab:ep-2015-processed} presents the walking episodes for the record
identification number \texttt{10041} from the TUS PUMF 2015 episode file
(Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
Only the unique active travel episode appears in Table
\ref{tab:ep-2015-processed} since the records were filtered to select
cases with walking or cycling episodes. For both cases, Tables
\ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain
labeled variables, facilitating the interpretation of the data.

\begin{table}
\centering
\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the 2015 TUS Main File.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{8}{10}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{cccccc}
\toprule
PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
\midrule
10000 & 616.6740 & July & 55 to 64 years & Male & Divorced\\
10001 & 8516.6140 & July & 55 to 64 years & Male & Married\\
10002 & 371.7520 & January & 45 to 54 years & Female & Married\\
10003 & 1019.3135 & March & 65 to 74 years & Female & Divorced\\
10004 & 1916.0708 & September & 25 to 34 years & Male & Single, never married\\
\addlinespace
10005 & 1952.2015 & April & 15 to 24 years & Male & Single, never married\\
10006 & 5761.5528 & August & 15 to 24 years & Male & Single, never married\\
10007 & 466.0426 & June & 55 to 64 years & Female & Widowed\\
10008 & 2479.2991 & February & 25 to 34 years & Female & Married\\
10009 & 1436.1641 & August & 65 to 74 years & Male & Widowed\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item Legend: PUMFID: record identification. WGHT\_PER:  person weight. SURVMNTH: survey month of data collection. AGEGR10: age group of the respondent. SEX: sex of the respondent. MARSTAT: marital status of the respondent.
\end{tablenotes}
\end{threeparttable}}
\end{table}

\begin{table}
\centering
\caption{\label{tab:gss-processed-file-2015}\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number 10041 of the 2015 GSS survey.}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{8}{10}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{rrlrlll}
\toprule
PUMFID & WGHT\_EPI & Activity & Duration & Origin & Destination & Mode\\
\midrule
10041 & 1353.818 & Transport to or from activity & 15 & Home & Home & Walking\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item Legend: PUMFID: record identification. EPINO: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
\end{tablenotes}
\end{threeparttable}}
\end{table}

@@

\begin{table}
\centering
\caption{\label{tab:table-01}\label{tab:table-01}Descriptive statistics of the duration (in minutes) of episodes with active transport records.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{>{}llccccccc}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{7}{c}{Year} \\
\cmidrule(l{3pt}r{3pt}){3-9}
Mode & Statistic & 1986 & 1992 & 1998 & 2005 & 2010 & 2015 & 2022\\
\midrule
 & Maximum & 660 & 300 & 255 & 515 & 480 & 900 & 480\\

 & Mean & 21 & 21 & 12 & 12 & 13 & 18 & 19\\

 & Median & 15 & 10 & 5 & 10 & 10 & 10 & 15\\

 & Minimum & 1 & 1 & 1 & 0 & 0 & 5 & 5\\

\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27 & 24\\
\cmidrule{1-9}
 & Maximum &  & 240 & 90 & 180 & 153 & 120 & 150\\

 & Mean &  & 28 & 24 & 20 & 19 & 25 & 40\\

 & Median &  & 15 & 15 & 15 & 10 & 20 & 30\\

 & Minimum &  & 5 & 2 & 1 & 1 & 5 & 5\\

\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20 & 20\\
\bottomrule
\end{tabular}
\end{table}
